---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(printr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# tidymargins

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/halpo/tidymargins.svg?branch=master)](https://travis-ci.org/halpo/tidymargins)
[![Test coverage](https://codecov.io/gh/halpo/tidymargins/branch/master/graph/badge.svg)](https://codecov.io/gh/halpo/tidymargins?branch=master)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![CRAN status](https://www.r-pkg.org/badges/version/tidymargins)](https://CRAN.R-project.org/package=tidymargins)
<!-- badges: end -->

The goal of `tidymargins` is to provide the `with_margins()` adverb 
function, as well as the related `spread_each()` function.


## Installation

You can install the released version of tidymargins from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("tidymargins")
```

## Basic Usage

`with_margins` is very simple to use.
It is an adverb, i.e. it accepts and returns a function.
The function, passed in, we'll call it `original`, should accept a 
[table like](https://tibble.tidyverse.org/) 
data object as it's first argument and return a similar object.
The resulting function, we'll call it `modified`, will also 
accept a data argument as it's first argument however that argument is expected to be a 
[grouped data frame](https://dplyr.tidyverse.org/reference/group_by.html). 
Other arguments to `modified` are passed along to the `original` function.
When `modified` is called with a grouped data frame `original` is called for 
every possible subset of grouping variables, including the full set 
of all original grouping variables and the empty set implying no grouping,
The results are bound together into a single data frame and returned.  
The example below shows this as an example.

```{r example}
library(dplyr)
library(tidymargins)
data(mtcars)

# our original function here is count
modified <- with_margins(count)

# modified is a function 
class(modified)

modified(group_by(mtcars, Cylinders = cyl, Gears = gear))
```

See that the output of modified is the counts grouped by; 
both `cyl` and `gear`, by `cyl` alone, by `gear` alone, and a 
grand total (no groups), all in a single data frame.  When a variable is 
not used as a grouping variable it is replaced with the value `"(All)`", 
indicating that all levels of that variable are included.  This label
can be changes by specifying the `all.name` argument when calling 
`with_margins`, shown in the next example.

There is no need to store the intermediate function returned by 
`with_margins`, but since the returned function may also accept other 
arguments care should be taken to what is passing to which function.
Care should also be taken with the pipe.

```{r}
mtcars %>% 
    select(cyl, gear, mpg, disp, hp, qsec) %>% 
    group_by(cyl, gear) %>% 
    # <---------- with_margins ----------------><-- summarise --->
    with_margins(summarise_if, all.name='Total')(is.numeric, mean) %>% 
    # making pretty
    rename( "Miles/(US) gallon" = "mpg"
          , "Cylinders" = "cyl"
          , "Displacement (cu.in.)" = "disp"
          , "Gross horsepower" = "hp"
          , "1/4 mile time" = "qsec"
          , "Number of forward gears" = "gear"
          )

```

